---
title: "Week 01 - Statistical Rethinking: A Bayesian Course"
subtitle: "Chapters 1, 2 and 3"
author: "Milan Wiedemann"
date: "`r lubridate::today()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rethinking)
# library(brms)
```

# Notes

## Chapter 1. The Golem of Prague

### Statistical golems

### Statistical rethinking

### Tools for golem engineering

## Chapter 2. Small Worlds and Large Worlds

### The garden of forking data

### Building a model

### Components of the model

### Making the model go

## Chapter 3. Sampling the Imaginary

### Sampling from a grid-approximate posterior

### Sampling to summarize

### Sampling to simulate prediction

# Homework

1. Suppose the globe tossing data (Chapter 2) had turned out to be 4 water in 15 tosses. 
Construct the posterior distribution, using grid approximation. 
Use the same flat prior as in the book.

```{r}

## Problem 1

# define grid
p_grid <- seq( from=0 , to=1 , length.out=1000 )

# define prior
prior <- rep( 1 , 1000 )

# compute likelihood at each value in grid
likelihood <- dbinom( 4 , size=15 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

# plot posterior
plot(x =  p_grid,
     y = posterior,
     type="l",
     xlab = "proportion water",
     ylab = "posterior probability" )
```

2. Start over in 1, but now use a prior that is zero below p = 0:5 and a constant above p = 0:5. 
This corresponds to prior information that a majority of the Earth's surface is water. 
What difference does the better prior make?

```{r}
## Problem 2

# define grid
p_grid <- seq( from=0 , to=1 , length.out=1000 )

# define prior
prior <- c( rep( 0 , 500 ) , rep( 2 , 500 ) )

# compute likelihood at each value in grid
likelihood <- dbinom( 4 , size=15 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

# plot posterior probability
plot(x = p_grid,
     y = posterior, 
     type = "l", 
     xlab = "proportion water", 
     ylab = "posterior probability")
```

3. For the posterior distribution from 2, compute 89% percentile and HPDI intervals. 
Compare the widths of these intervals. 
Which is wider? Why? 
If you had only the information in the interval, what might you misunderstand about the shape of the posterior distribution?

```{r}
## Problem 3

samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )

plot( samples , ylim=c(0,1) , xlab="samples" , ylab="proportion water" )

PI( samples , 0.89 )

HPDI( samples , 0.89)

```

