---
title: "Week 1"
author: "Chris Beeley"
date: "08/03/2021"
output:
  html_document: default
  pdf_document: default
---
  
```{r setup, include=FALSE}

library(rethinking)

knitr::opts_chunk$set(echo = TRUE)

```

## Chapter 4

### E1

In the model below, which line is the likelihood:

$$y_{i} \sim Normal(\mu, \sigma)$$
$$\mu \sim Normal(0, 10)$$
$$\sigma \sim Exponential(1)$$

Top line (the other lines are priors).

### E2

In the model definition above, how many parameters are in the posterior distribution?

Two. Mu and sigma

### E3

Using the model definition above, write down the appropriate form of Bayes' theorem that includes the proper likelihood and priors.

$$Pr(\mu, \sigma | y_{i}) = \frac{Normal(y_{1} | \mu, \sigma).Normal(p |0, 10)}{\int Normal(y_{1} | \mu, \sigma).Normal(p| 0, 10) d\mu}$$

I have no clue, to be honest

### E4

In the model definition below, which line is the linear model?

$$y_{i} \sim Normal(\mu, \sigma)$$
$$\mu_{i} = \alpha + \beta x_{i}$$
$$\alpha \sim Normal(0, 10)$$
$$\beta \sim Normal(0, 1)$$
$$\sigma \sim Exponential(2)$$

Line two is the linear model.

### E5

In the model above, how many parameters are in the posterior distribution? Three- $\alpha$, $\beta$, and $\sigma$.

### M1

For the following model, simulate observed values from the prior:

$$y_{i} \sim Normal(\mu, \sigma)$$
$$\mu \sim Normal(0, 10)$$
$$\sigma \sim Exponential(1)$$
Plot it first: 

```{r}

par(mfrow = c(2, 1))

curve(dnorm(x, 0, 10), from = -50, to = 50, main = "mean")
curve(dexp(x, 1), from = 0, to = 1, main = "sd")

par(mfrow = c(1, 1))
```

Now sample from it

```{r}

sample_mu <- rnorm(1e4, 0, 10)
sample_sigma <- rexp(1e4, 1)

prior <- rnorm(1e4, sample_mu, sample_sigma)
dens(prior)

```

### M2

Translate the model above into a quap formula

```{r}

flist <- alist(
  value ~ dnorm(mu, sigma),
  mu ~ dnorm(0, 10),
  sigma ~ dexp(1)
)

```

Translate the following code into a mathematical formula

```{r, eval = FALSE}

y ~ dnorm(mu, sigma)
mu <- a + b * x
a ~ dnorm(0, 10)
b ~ dunif(0, 1)
sigma ~ dexp(1)

```

$$y_{i} \sim Normal(\mu, \sigma)$$
$$\mu = a + \beta x$$
$$a \sim Normal(0, 10)$$
$$b \sim Uniform(0, 1)$$
$$\sigma \sim Exponential(1)$$


